### 2013-06-08 doc/pgas13/evaluation.tex
Related work and citations galore.
Pretty minor wording and citation fixes

\section{Evaluation}
To evaluate the impact flat combining has on the global data structures, we ran a series of experiments to test the raw performance of the data structures themselves under different workloads, and further measured their impact on performance of two simple graph benchmarks.

@@

Figure~\ref{fig:vector} shows the results of the throughput experiments for the global Stack and Queue. Results are shown with flat combining completely disabled, only combining at the master core, only combining locally, and full two-level combining.

Despite Grappa's automatic aggregation, without combining, both the stack and queue completely fail to scale because all workers' updates must serialize on the master core.
Though combining on the master core alleviates some of the serialization, it does not provide significant benefit because all of the synchronization must still be performed on one core. However, with local flat combining, synchronization is able to be done mostly in parallel, with much less-frequent bulk synchronization at the master.

On the mixed workload, the stack is able to do matching locally, allowing it to reduce the amount of communication drastically, greatly improving its performance. Figure~\ref{fig:stack_stats} corroborates this, showing that the amount of combining that occurs directly correlates with the observed throughput.

We also experimented with performing both forms of combining together; however, combining at the master did not affect performance significantly. This was somewhat surprising as combined updates will still force others to wait. However, with both levels of combining enabled, relatively few requests are combined at the master---for the all-push workload, on the order of a thousand, and for the mixed workload on the stack, the average is less than 20.

The queue benefits in the same way from reducing synchronization and batching, and its all-push workload performs identically to the stack's.
However, the queue is unable to do matching locally, and in fact, the mixed workload performs worse because the current implementation serializes combined pushes and combined pops. This restriction could be lifted with more careful synchronization at the master core allowing pushes and pops to proceed in parallel as long as there are enough elements in the queue.

@@

Without combining, performance scales out to 32 nodes because synchronization is distributed over hash cells, but drops off as the number of destinations increases.

Figure~\ref{fig:hash_perf} shows the throughput results for the GlobalHashSet and GlobalHashMap.
Both data structures synchronize at each hash cell, which allows them to scale fairly well even without combining. However, after 32 nodes, scaling drops off significantly due to increased number of concurrent accessors and destinations.
Combining allows duplicate inserts and lookups to be eliminated, so performs better the smaller the key range. This reduction in message traffic allows scaling to continue out to 64 nodes.

\subsection{Application Kernel Performance}
The original goal was to have scalable global data structures in the Grappa library that could be used in applications. Naive implementations of the data structures are insufficient for scaling, so HPC applications often implement their own highly customized versions. However, this forces each application to carefully reason about the ways in which the structures will be accessed. Additionally, many optimizations rely on relaxing global consistency in a way that does not affect the program. 
The Grappa data structures are synchronized to provide the most general use and match the expectations of programmers and algorithms.

\paragraph{Breadth-First Search}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{data/plots/bfs_perf.pdf}
  \caption{\emph{BFS} on a Graph500-spec graph of scale 26 (64 million vertices, 1 billion edges), with the direction-optimizing BFS algorithm. Performance is measured in millions of Traversed Edges Per Second (MTEPS).}
  \label{fig:bfs_perf}
\end{figure}
The first application kernel is the Graph500 Breadth-First-Search (BFS) benchmark. This benchmark does a search starting from a random vertex in a synthetic graph and builds a search tree of parent vertices for each vertex traversed during the search. While this is a relatively simple problem, it exercises the random-access throughput of a system as well as being a primitive in many other graph algorithms. The BFS algorithm contains a global queue which represents the frontier of vertices to be visited in each level.
Our implementation employs the direction-optimizing algorithm by Beamer et al.\cite{Beamer:Graph500} which performs particularly well for the scale-free RMAT graphs generated by the benchmark.
The frontier queue in BFS is amenable to further optimization to take advantage of the fact that the algorithm does pushes and pops in separate phases, allowing consistency to be relaxed.
We compare our implementation of BFS using the flat-combined global queue described above with a highly tuned Grappa implementation that uses a custom asynchronous queue.

Figure~\ref{fig:bfs_perf} shows the results of scaling the BFS kernel up to 64 nodes. The simple queue implementation without flat combining is completely unscalable. However, the flat combining queue tracks the highly tuned asynchronous version. This illustrates that providing a safe, synchronized data structure for initially developing algorithms for PGAS is possible, while further optimizations can be applied incrementally.

\paragraph{Connected Components}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{data/plots/cc_perf.pdf}
  \caption{\emph{Connected Components} on the same scale 26 Graph500 graph. Performance is measured in MTEPS.}
  \label{fig:cc_perf}
\end{figure}
Connected Components (CC) is another simple graph analysis kernel that illustrates another use of global data structures in irregular applications. We implement the three-phase CC algorithm~\cite{mtgl} which was designed for the massively-parallel MTA-2 machine. In first phase, the algorithm begins many traversals in parallel from random starting vertices, labeling vertices with the root vertex. Whenever two traversals encounter each other, their searches are pruned and an edge between the two roots is inserted in a set. After all edges have been traversed in this way, the set of edges forms a new, typically much smaller, graph. The second phase performs the classical Shiloach-Vishkin parallel algorithm\cite{shiloach1982n} on this reduced graph, and the final phase propagates the component labels out to the full graph.
The creation of the reduced-graph edge set dominates the runtime of this algorithm, so improving the implementation of the set operations has a significant impact on performance. As in the case of BFS, further optimizations involving relaxation of consistency can be applied, in this case, to the global set. Therefore, we compare our straightforward implementation using the generic GlobalHashSet with and without flat combining against a tuned asynchronous implementation.

The results in Figure~\ref{fig::cc_perf} show that none of these three implementations scale very well out to 64 nodes. However, performing combining does improves the performance of the algorithm overall and improves scaling. The tuned version outperforms the synchronous version because it is able to build up most of the set locally on each core before merging them at the end of the first phase. An implementation of the GlobalHashSet that did not provide synchronized semantics could potentially relax consistency in a more general way, but this is left for future work.

>>>>

\section{Evaluation}
To evaluate the impact flat combining has on the global data structures, we ran a series of experiments to test the raw performance of the data structures themselves under different workloads, and then measure their impact on performance of two simple graph benchmarks.

@@

Figure~\ref{fig:vector} shows the results of the throughput experiments for the global Stack and Queue. Results are shown with flat combining completely disabled, only combining at the master core (``centralize''), and combining locally (``distributed'').

Despite Grappa's automatic aggregation, without combining, both the stack and queue completely fail to scale because all workers' updates must serialize.
Though centralized combining alleviates some of the serialization, its benefit is limited because all synchronization must still be performed on one core. However, with local flat combining, synchronization is able to be done mostly in parallel, with less-frequent bulk synchronization at the master.

On the mixed workload, the stack is able to do matching locally, allowing it to reduce the amount of communication drastically, greatly improving its performance on the mixed workload. Figure~\ref{fig:stack_stats} corroborates this, showing that the amount of combining that occurs directly correlates with the observed throughput.

We also experimented with performing both forms of combining together; however, combining at the master did not affect performance significantly. This was somewhat surprising as combined updates still force others to wait. However, with both levels of combining enabled, relatively few requests are combined at the master---for the all-push workload, on the order of a thousand, and for the mixed workload on the stack, the average is less than 20.

The queue benefits in the same way from reducing synchronization and batching, and its all-push workload performs identically to the stack's.
However, the queue is unable to do matching locally, and in fact, the mixed workload performs worse because the current implementation serializes combined pushes and combined pops. This restriction could be lifted with more careful synchronization at the master core.

@@

% Without combining, performance scales out to 32 nodes because synchronization is distributed over hash cells, but drops off as the number of destinations increases.
Figure~\ref{fig:hash_perf} shows the throughput results for the Set and Map.
Both data structures synchronize at each hash cell, which allows them to scale fairly well even without combining. However, after 32 nodes, scaling drops off significantly due to the increased number of destinations.
Combining allows duplicate inserts and lookups to be eliminated, so performs better the smaller the key range. This reduction in message traffic allows scaling to continue out to 64 nodes.

\subsection{Application Kernel Performance}
The original goal was to have scalable global data structures in the Grappa library that could be used in applications. Generic implementations of data structures are often insufficient for scaling, so HPC applications often implement their own customized versions. However, this forces each application to carefully reason about the ways in which the structures will be accessed, and often relies on relaxing consistency.
The Grappa data structures are synchronized to provide the most general use and match the expectations of programmers and algorithms.
In these evaluations we compare the flat-combining structures against tuned, relaxed-consistency versions.

\paragraph{Breadth-First Search}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{data/plots/bfs_perf.pdf}
  \caption{\emph{BFS} on a Graph500-spec graph of scale 26 (64 million vertices, 1 billion edges), with the direction-optimizing BFS algorithm. Performance is measured in millions of Traversed Edges Per Second (MTEPS).}
  \label{fig:bfs_perf}
\end{figure}
The first application kernel is the Graph500 Breadth-First-Search (BFS) benchmark~\cite{graph500list}. This benchmark does a search starting from a random vertex in a synthetic graph and builds a search tree of parent vertices for each vertex traversed during the search. While this is a relatively simple problem, it exercises the random-access throughput of a system as well as being a primitive in many other graph algorithms. The BFS algorithm contains a global queue which represents the frontier of vertices to be visited in each level.
Our implementation employs the direction-optimizing algorithm by Beamer et al.~\cite{Beamer:Graph500} which performs particularly well for the scale-free RMAT graphs generated by the benchmark.
The frontier queue in BFS is amenable to further optimization to take advantage of the fact that the algorithm does pushes and pops in separate phases, allowing consistency to be relaxed.
We compare our implementation of BFS using the flat-combined global queue described above with a highly tuned Grappa implementation that uses a custom asynchronous queue.

Figure~\ref{fig:bfs_perf} shows the results of scaling the BFS kernel up to 64 nodes. The simple queue implementation without flat combining is completely unscalable. However, the flat combining queue tracks the tuned version. This illustrates that providing a safe, synchronized data structure for initially developing algorithms for PGAS is possible without sacrificing scaling.
% , while further optimizations can be applied incrementally.

\paragraph{Connected Components}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{data/plots/cc_perf.pdf}
  \caption{\emph{Connected Components} on the same scale 26 Graph500 graph. Performance is measured in MTEPS.}
  \label{fig:cc_perf}
\end{figure}
Connected Components (CC) is another simple graph analysis kernel that illustrates a different use of global data structures in irregular applications. We implement the three-phase CC algorithm~\cite{mtgl} which was designed for the massively-parallel MTA-2 machine. In the first phase, parallel traversals attempt to claim vertices and label them.
Whenever two traversals encounter each other, an edge between the two roots is inserted in a set. After all edges have been traversed in this way, the set of edges forms a new, typically much smaller, graph. The second phase performs the classical Shiloach-Vishkin parallel algorithm~\cite{shiloach1982n} on this reduced graph, and the final phase propagates the component labels out to the full graph.
The creation of the reduced-graph dominates the runtime of this algorithm, in includes many repeated inserts at the boundary between traversals.
As in BFS, further optimizations involving relaxation of consistency can be applied, in this case to the Set. Therefore, we compare our straightforward implementation using the generic Set with and without flat combining against a tuned asynchronous implementation.

The results in Figure~\ref{fig:cc_perf} show that none of these three scale well out to 64 nodes. However, performing combining does improve the performance over the uncombined case. The tuned version outperforms the synchronous version because it is able to build up most of the set locally on each core before merging them at the end of the first phase. An implementation that did not provide synchronized semantics could potentially relax consistency in a more general way, but this is left for future work.

### 2013-06-08 doc/pgas13/flat-combining.tex
Pretty minor wording and citation fixes

% In most cases, this means communicating with whichever core holds the ``lock-protected'' field that maintains sequential consistency. In the case of a global stack, this is the ``master'' \emph{top} pointer which must be incremented on pushes and decremented on pops. On the core where global synchronization is occurring, there must be synchronization between the multiple synchronization requests arriving from other cores.

% At this point we have traded off one remote synchronization per operation for an additional local synchronization per operation and some number fewer remote synchronizations (depending on the amount of combining that occurs). This would likely result in a net benefit due to local synchronization being inherently cheaper. However, because of the way the Grappa runtime schedules computation, it is actually possible to get the correct mutual exclusion/atomicity without any additional synchronization overhead.

% By design, each core in Grappa operates independently of the other cores. Workers are scheduled cooperatively, so each worker can assume atomicity until it performs an operation that yields to the scheduler, such as a remote call. Therefore workers within a core can cooperate without any explicit synchronization operations requiring memory fences. This allows them to publish requests to the local proxy very inexpensively, eliminating the need for a complex synchronization-amortizing queue. This has the added benefit that there will be no unused publication records, which could have become a problem with thousands of workers. Further, remote operations are processed through the same cooperative-multithreading scheduler on the remote side, so combined operations are trivially serialized and atomic.

\paragraph{Memory Consistency Model}

In the context of this work, sequential consistency guarantees that within a
task, operations will be in task order and that all tasks will observe the
same global order of operations. An example is shown in
Figure~\ref{fig:sync}.\TODO{make sync example}

>>>>

% In most cases, this means communicating with whichever core holds the ``lock-protected'' field that maintains sequential consistency. In the case of a global stack, this is the ``master'' \emph{top} pointer which must be incremented on pushes and decremented on pops. On the core where global synchronization is occurring, there must be synchronization between the multiple synchronization requests arriving from other cores.
% At this point we have traded off one remote synchronization per operation for an additional local synchronization per operation and some number fewer remote synchronizations (depending on the amount of combining that occurs). This would likely result in a net benefit due to local synchronization being inherently cheaper. However, because of the way the Grappa runtime schedules computation, it is actually possible to get the correct mutual exclusion/atomicity without any additional synchronization overhead.
% By design, each core in Grappa operates independently of the other cores. Workers are scheduled cooperatively, so each worker can assume atomicity until it performs an operation that yields to the scheduler, such as a remote call. Therefore workers within a core can cooperate without any explicit synchronization operations requiring memory fences. This allows them to publish requests to the local proxy very inexpensively, eliminating the need for a complex synchronization-amortizing queue. This has the added benefit that there will be no unused publication records, which could have become a problem with thousands of workers. Further, remote operations are processed through the same cooperative-multithreading scheduler on the remote side, so combined operations are trivially serialized and atomic.

\paragraph{Memory Consistency Model}
In the context of this work, sequential consistency guarantees that within a
task, operations will be in task order and that all tasks will observe the
same global order of operations.
% An example is shown in Figure~\ref{fig:sync}.\TODO{make sync example}

### 2013-06-08 doc/pgas13/grappa.tex
Pretty minor wording and citation fixes

Irregular applications are characterized by having unpredictable data-dependent access patterns and poor spatial and temporal locality. Applications in this class, including data mining, graph analytics, and various learning algorithms, are becoming increasingly prevalent in high-performance computing. These programs typically perform many fine-grained accesses to disparate sources of data, which is a problem even at multicore scales, but is further exacerbated on distributed memory machines.
It is often the case that naively porting an application to a PGAS system results in excessive communication and poor access patterns, but this class of applications defies typical optimization techniques such as data partitioning, shadow objects, and bulk-synchronous communication transformations.
Luckily, applications in this class have another thing in common: abundant amounts of data parallelism. This parallelism can be exploited in a number of different ways to improve overall throughput of these applications.

@@

Grappa is a PGAS-style runtime for commodity clusters which has been designed from the ground up to achieve high performance on irregular applications. The key is latency tolerance---long-latency operations such as remote communication can be tolerated by switching to another concurrent thread of execution. Given abundant  concurrency, there are opportunities to increase throughput by sacrificing latency. In particular, throughput of random accesses to remote memory can be improved by delaying communication requests and aggregating them into larger packets. Highly-tuned implementations of irregular applications in shared-memory, PGAS and message-passing paradigms, typically end up implementing constructs to allow for aggregation and leverage data parallelism. The goal of Grappa is to provide a simple-to-use PGAS abstraction with mechanisms for expressing the concurrency needed to achieve high performance.

@@

\subsection{Aggregated Communication}
Communication in Grappa is achieved via \emph{active messages} in conjunction with bulk remote direct memory access (RDMA) operations. Various components of Grappa's runtime contribute different kinds of messages to move and modify remote data, perform synchronization, or spawn tasks. To make efficient use of the networks in high-performance systems, which perform better with larger messages (typically 64 KB), all communication in Grappa is sent through an aggregation layer that automatically accumulates messages to the same destination. When enough messages are ready to go to a particular destination node, they are serialized into a buffer and sent using RDMA.

\subsection{Global Memory}
In the PGAS style, Grappa has a notion of a global address space which is partitioned across the physical memories of the nodes in a cluster. Each core owns a portion of memory, which is further divided for three different uses:
\begin{itemize}
  \item Execution stacks for the core's Workers.
  \item A core-local heap.
  \item A slice of the global shared heap.
\end{itemize}
All of these can be addressed by any core in the system using a Grappa \texttt{GlobalAddress}, which encodes both the owning core and the address on that core. Additionally, addresses into the global heap are partitioned in a block-cyclic fashion, so that a large allocation is automatically distributed among many nodes. For irregular applications, this helps avoid hot spots and is typically sufficient for random access.

All accesses to memory owned by another core must be done by that core via a message. In this way, the runtime maintains strict isolation even between processes running on the same physical memory. At the programming level, however, this is hidden behind higher-level remote operations, which in Grappa are called ``delegates''. Figure~\ref{fig:system} shows an example of a delegate \texttt{read()} which blocks the calling task and sends a message to the owner, who sends a reply with the data and wakes the caller.

>>>>

Irregular applications are characterized by having unpredictable data-dependent access patterns and poor spatial and temporal locality. Applications in this class, including data mining, graph analytics, and various learning algorithms, are becoming increasingly prevalent in high-performance computing. These programs typically perform many fine-grained accesses to disparate sources of data, which is a problem even at multicore scales, but is further exacerbated on distributed memory machines.
It is often the case that naively porting an application to a PGAS system results in excessive communication and poor access patterns~\cite{pgas-analysis}, but this class of applications defies typical optimization techniques such as data partitioning, shadow objects, and bulk-synchronous communication transformations.
Luckily, applications in this class have another thing in common: abundant amounts of data parallelism. This parallelism can be exploited in a number of different ways to improve overall throughput of these applications.

@@

Grappa is a global-view PGAS-style runtime for commodity clusters which has been designed from the ground up to achieve high performance on irregular applications. The key is latency tolerance---long-latency operations such as remote communication can be tolerated by switching to another concurrent thread of execution. Given abundant  concurrency, there are opportunities to increase throughput by sacrificing latency. In particular, throughput of random accesses to remote memory can be improved by delaying communication requests and aggregating them into larger packets. Highly-tuned implementations of irregular applications in shared-memory, PGAS and message-passing paradigms, typically end up implementing constructs to allow for aggregation and leverage data parallelism. The goal of Grappa is to provide a simple-to-use PGAS abstraction with mechanisms for expressing the concurrency needed to achieve high performance.

@@

\subsection{Aggregated Communication}
Communication in Grappa is achieved via \emph{active messages}~\cite{vonEicken92} in conjunction with bulk remote direct memory access (RDMA) operations. Various components of Grappa's runtime contribute different kinds of messages to move and modify remote data, perform synchronization, or spawn tasks. To make efficient use of the networks in high-performance systems, which perform better with larger messages (typically 64 KB), all communication in Grappa is sent through an aggregation layer that automatically accumulates messages to the same destination. When enough messages are ready to go to a particular destination node, they are serialized into a buffer and sent using RDMA.

\subsection{Global Memory}
In the PGAS style, Grappa has a notion of a global address space which is partitioned across the physical memories of the nodes in a cluster. Each core owns a portion of memory, which is divided among execution stacks for the core's Workers, a core-local heap, and a slice of the global heap.

All of these can be addressed by any core in the system using a Grappa \texttt{GlobalAddress}, which encodes both the owning core and the address on that core. Additionally, addresses into the global heap are partitioned in a block-cyclic fashion, so that a large allocation is automatically distributed among many nodes. For irregular applications, this helps avoid hot spots and is typically sufficient for random access.

All accesses to memory owned by another core must be done by that core via a message. In this way, the runtime maintains strict isolation even between processes running on the same physical memory. At the programming level, however, this is hidden behind higher-level remote operations, which in Grappa are called ``delegate'' operations. Figure~\ref{fig:system} shows an example of a delegate \texttt{read()} which blocks the calling task and sends a message to the owner, who sends a reply with the data and wakes the caller.

### 2013-06-08 doc/pgas13/intro.tex
Pretty minor wording and citation fixes

The goal of partitioned global address space (PGAS)~\cite{upc:2005} languages and runtimes is to provide the illusion of a single shared memory to a program that actually executes on a number of machines each with their own memory. This allows programmers to write their algorithms without needing to explicitly manage communication. Once the algorithm is correct, many techniques exist to help improve performance improving locality (spatial and temporal) and coarsening communication where possible. However, with this shared memory abstraction come all of the difficult concurrency issues that arise in physically-shared memory. Luckily, there exists a large body of work solving these issues in physically-shared memory which the PGAS community can leverage. The opportunity here is that, in a distributed setting, the costs of communication and synchronization operations are different, so different performance trade-offs will be made.

@@

The trick is coming up with a mechanism for delegating that has lower synchronization overhead than the original contention case.
Hendler et al.\cite{flatCombining} showed that a cleverly-implemented ``publication list'' can allow multiple accessors to cooperate with minimal synchronization.
The same work makes the additional observation that, given the semantics of some particular data structures, the set of accesses being serialized can be composed and performed more efficiently \emph{combined} than individually, which they dub \emph{flat combining}.
Together, the reduced synchronization cost and combined operations allow even a data structure with a single global lock to scale better than complicated concurrent data structure implementations with fine-grained locking.
Further, the same publication list mechanism can be applied to other data structures, and all that must be customized is the particular way in which operations for a given data structure can be combined.

@@

The goal of this work is to apply flat-combining paradigm to a PGAS runtime to reduce the cost of maintaining sequentially consistent data structures.
We leverage the Grappa runtime~\cite{Nelson:hotpar11-real}, a PGAS-style runtime library optimized for fine-grained random access, which provides the ability to tolerate long latencies by efficiently switching between many lightweight threads.
By leveraging lots of concurrency, we find that it is actually possible to scalably enforce strict consistency among threads. We leverage Grappa's latency tolerance mechanisms to allow many fine-grained synchronized operations to be combined to reduce communication and achieve higher overall throughput.
In addition, we show how a generic flat-combining framework can be used to implement multiple scalable sequentially-consistent global data structures.

% lay out the upcoming sections of the paper...
The next section will describe in more detail the Grappa runtime system that is used to implement flat combining for distributed memory machines. Then we explain the flat combining paradigm in more depth and describe how it maps to a PGAS model. Next we explain how several data structures are implemented in our framework and show how they perform on simple throughput workloads. Finally, we evaluate how this affects application performance in two simple graph analysis kernels.

>>>>

The goal of partitioned global address space (PGAS)~\cite{upc:2005} languages and runtimes is to provide the illusion of a single shared memory to a program that actually executes on a number of machines each with their own memory. This allows programmers to write their algorithms without needing to explicitly manage communication. Once the algorithm is correct, many techniques exist to help improve performance improving locality (spatial and temporal) and coarsening communication where possible. However, with this shared memory abstraction come all of the difficult concurrency issues that arise in physically-shared memory. Luckily, there exists a large body of work solving these issues in physically-shared memory which the PGAS community can leverage. In a distributed setting, the costs of communication and synchronization operations are different, providing an opportunity for findings in shared-memory to be re-evaluated.

@@

The trick is coming up with a mechanism for delegating that has lower synchronization overhead than the original contention case.
Hendler et al.\cite{flatCombining} showed that a cleverly-implemented ``publication list'' can allow multiple accessors to cooperate with minimal synchronization cost.
In addition, given the semantics of some particular data structures, the set of accesses being serialized can be composed and performed more efficiently \emph{combined} than individually.
Together, the reduced synchronization cost and combined operations allow even a data structure with a single global lock to scale better than complicated concurrent data structures using fine-grained locking or lock-free mechanisms.
The same publication list mechanism can be applied to multiple data structures; all that must be customized is the particular way in which the data structure's operations can be combined.

@@

The goal of this work is to apply the flat-combining paradigm to a PGAS runtime to reduce the cost of maintaining sequentially consistent data structures.
We leverage the Grappa runtime~\cite{Nelson:hotpar11-real}, a PGAS-style runtime library optimized for fine-grained random access, which provides the ability to tolerate long latencies by efficiently switching between many lightweight threads.
By leveraging lots of concurrency, we find that it is actually possible to scalably enforce strict sequential consistency. We leverage Grappa's latency tolerance mechanisms to allow many fine-grained synchronized operations to be combined to reduce communication and achieve higher overall throughput.
In addition, we show how a generic flat-combining framework can be used to implement multiple global data structures.

% lay out the upcoming sections of the paper...
The next section will describe in more detail the Grappa runtime system that is used to implement flat combining for distributed memory machines. We then explain the flat combining paradigm in more depth and describe how it maps to a PGAS model. Next we explain how several data structures are implemented in our framework and show how they perform on simple throughput workloads. Finally, we evaluate how this affects application performance in two simple graph analysis kernels.

### 2013-06-08 doc/pgas13/relatedwork.tex
Pretty minor wording and citation fixes

\section{Related Work}
Attempts to improve the scalability of synchronized data structures are as old as the data structures themselves. Though much attention has gone to lock-free data structures such as the Trieber Stack~\cite{treiber} and Michael-Scott Queue~\cite{msqueue}, a good body of work has explored ways of scaling globally-locked data structures.

>>>>

\section{Related Work}
Attempts to improve the scalability of synchronized data structures are as old as the data structures themselves. Though much attention has gone to lock-free data structures such as the Treiber Stack~\cite{treiber} and Michael-Scott Queue~\cite{msqueue}, a good body of work has explored ways of scaling globally-locked data structures.
